### 2024-10-16 Article/main.tex
prove

In the category of graphs with equivalences, universal constructions are easy to obtain. 
\begin{prop}\label{prop:limits_in_EqGrph}
    COME SI CALCOLANO I LIMITI IN $\mathbf{EqGrph}$
    % GUESS: limiti calcolati componente per componente su E e V, e restrizioni su Q (?) come in Graph (Dipende da prefasci).
\end{prop}
\begin{proof}
    TO DO
\end{proof}
 \todo{riscrivere questa parte come esempi che derivano dalla proposizione precedente}
The terminal object in $\mathbf{EqGrph}$ is the graph with only one node, only one edge and with the single node equivalent only to itself, while the initial object is the graph with no nodes and no edges.
Given two graphs with equivalences $\eqgraph{G} = (G, \sim_G) $ and $\eqgraph{H} = (H, \sim_{H})$, one can construct the product graph $\eqgraph{G \times H} = (G \times H, \sim_{G \times H})$, together with the two projections $\pi_G$ and $\pi_H$,  where $\sim_{G \times H}$ is the least equivalence relation on $G \times H$ such that, if $w_1$ and $w_2$ are nodes of $\eqgraph{G \times H}$, we have $w_1 \sim_{G \times H} w_2$ if $\pi_G(w_1) \sim_G \pi_G(w_2)$ and $\pi_H(w_1) \sim_H \pi_H(w_2)$.
An equalizer in $\mathbf{EqGrph}$ for a pair of morphisms $f, g: \eqgraph{G \rightarrow H}$ is exactly the same as in $\mathbf{Graph}$, namely, $(\eqgraph{E} = (E, \sim_E), e)$, where $E$ is the greatest subgraph of the underlying graph of $\eqgraph{G}$ on which $f$ and $g$ are equal, together with the inclusion $e: \eqgraph{E \rightarrow G}$. The equivalence relation is preserved ($\sim_E$ is $\sim_G$ restricted to the set of vertices of $E$).

@@

    %%% IDEA: far vedere che, poiché i colimiti in una categoria di prefasci sono calcolati punto punto, allora ho un pushout su ogni componente (E, V, Q). Dalla prop nella parte sull'adesività si può concludere che tutti e tre i pushout preservano mono e quindi mono è preservata su tutte e tre le componenti, quindi reg  mono.
\end{proof}
}\color{black}
\todo[color=green!40]{
COSE DA FARE
>Quoziente crea i limiti (questo dovrebbe derivare direttamente dalla sezione precedente ) 
>Quoziente crea certi pushout (questo va fatto ex novo)
}
% Regular monos in $\mathbf{EqGrph}$ are then natural transformations $\eta: \eqgraph{G \dot\rightarrow H}$ such that $\eta_Q$ is mono (\Cref{prop:reg_mono_in_EG_are_mono_in_graph}). Pullbacks preserve regular monomorphisms (\Cref{prop:monos_pres_by_pullback}). Regular monos are preserved also by pushouts (\Cref{lemma:reg_mono_pres_by_pushouts}). Again, since (co)limits are computed pointwise, and in $\Set$ monos are preserved by pushouts (since $\Set$ is adhesive), we can conclude what follows.

@@

\begin{proof}
    We will show now that $Q$ reflects pullbacks.
    Consider the following diagram in $\mathbf{Graph}$
    \[
        \begin{tikzcd}
            & Q(\eqgraph{G}) \arrow[d, "Q(f)"] \\
            Q(\eqgraph{H}) \arrow[r, "Q(g)"swap] & Q(\eqgraph{K})
        \end{tikzcd}
    \]
    and let $(\mathcal P, p_{Q(f)}, p_{Q(g)})$ be the pullback. We want to show that, if  $(\eqgraph{P}, p_f, p_g)$ is the pullback of the arrows $f$ and $g$ in $\mathbf{EqGrph}$, then $Q(\eqgraph{P}) = \mathcal{P}$, $Q(p_f) = p_{Q(f)}$ and $Q(p_g) = p_{Q(g)}$. 
    \color{green}{
        Since limits are computed pointwise in $\mathbf{EqGrph}$ on $E$ and $V$ components, we have that 
        % IDEA: come si calcola un pullback in EgGrph? --> componente per componente, e poi su Q la least equiv relation such that ..., quindi applicando il funtore quoziente ottengo esattamente il pullback dei quozienti.
    }\color{black}
    Hence, from \Cref{cor:quot_preserves_co_lim}, we can conclude that $Q$ creates pullbacks.
\end{proof}

@@

\begin{proof}
    1) --> Q preserva pushout along reg monos ()
\end{proof}

>>>>

In the category of graphs with equivalences, universal constructions are easy to obtain. 
\begin{prop}\label{prop:limits_in_EqGrph}
    Let $D: \cat I \rightarrow \mathbf{EqGrph}$ be a diagram. Then, a limit for $D$ is $(\eqgraph{L}, l^i)_{i\in \Ob(\cat I)}$ with in which $\eqgraph{L} = (\mathcal{L}, \sim_{\mathcal{L}})$ is a graph in which (in $\Set$):
    \begin{itemize}
        \item $(V_{\mathcal{L}}, l^i_{V})$ is the limit of all the $V_{D(i)}$;
        \item $(E_{\mathcal{L}}, l^i_{E})$ is the limit of all the $E_{D(i)}$.
    \end{itemize}
    And the equivalence relation on nodes is the least equivalence relation $\sim_{\mathcal{L}}$ such that, given two vertices $v$ and $w$ of $\eqgraph{L}$, if $v\sim_{\mathcal{L}}w$m then $l^i_V(v) \sim_i l^i_V(w)$ for each $i\in \Ob(\cat I)$. (Here $\sim_i$ denotes the equivalence relation of $D(i)$).
\end{prop}

\begin{proof}

    \color{red}{
    Firstly, we show that $(\eqgraph{L}, l^i)_{\cat I}$ is a cone. Let $D(\alpha) : D(i) \rightarrow D(j)$, with $\alpha: i \rightarrow j$ morphism in $\cat I$. Our claim is $D(\alpha) \circ l^i = l^j$. Hence, 
    \begin{align*}
        D(\alpha) \circ l^i &= (h_V, h_E) \circ (l^i_V, l^i_E)      \\
                            &= (h_V \circ l^i_V, h_E \circ l^i_E)   \\
                            &= (l^j_V, l^j_E)                       \\
                            &= l^j                              
    \end{align*}
    For $(\eqgraph{L}, l^i)_{\cat I}$ to be a cone, the only constraint on the equivalence relation $\sim_{\mathcal{L}}$ is to be such that graph homomorphisms above can preserve it. But such equivalence relation exists in every case, considering the identity. This led us to the second point of the proof, that is, the universal property. 
    % L'idea qui è di prendere la relazione di equivalenza tale che il grafo limite abbia la proprietà universale.
    For the cone described above to be a limit must be valid the universal property. Such condition is is satisfied taking the least equivalence relation such that, given two vertices $v$ and $w$ of $\eqgraph{L}$, if $v\sim_{\mathcal{L}}w$ then $l^i_V(v) \sim_i l^i_V(w)$ for each $i\in \Ob(\cat I)$. Suppose the following situation:
    \[
        \begin{tikzcd}
            & \eqgraph{L}'  \arrow[d, "h"] \arrow[ddl, bend right=30, "{q^i}"swap] \arrow[ddr, bend left=30, "{q^j}"] & \\
            & \eqgraph{L} \arrow[dl, "{l^i}"swap] \arrow[dr, "{l^j}"] & \\
            D(i) \arrow[rr, "{D(\alpha)}"]& &  D(j)
        \end{tikzcd}
    \]
    The universal property the follows from the definition of least equivalence relation.
    } \color{black}
\end{proof}

 \todo{riscrivere questa parte come esempi che derivano dalla proposizione precedente}
The terminal object in $\mathbf{EqGrph}$ is the graph with only one node, only one edge and with the single node equivalent only to itself, while the initial object is the graph with no nodes and no edges.
Given two graphs with equivalences $\eqgraph{G} = (G, \sim_G) $ and $\eqgraph{H} = (H, \sim_{H})$, one can construct the product graph $\eqgraph{G \times H} = (G \times H, \sim_{G \times H})$, together with the two projections $\pi_G$ and $\pi_H$,  where $\sim_{G \times H}$ is the least equivalence relation on $G \times H$ such that, if $w_1$ and $w_2$ are nodes of $\eqgraph{G \times H}$, we have $w_1 \sim_{G \times H} w_2$ if $\pi_G(w_1) \sim_G \pi_G(w_2)$ and $\pi_H(w_1) \sim_H \pi_H(w_2)$.
An equalizer in $\mathbf{EqGrph}$ for a pair of morphisms $f, g: \eqgraph{G \rightarrow H}$ is exactly the same as in $\mathbf{Graph}$, namely, $(\eqgraph{E} = (E, \sim_E), e)$, where $E$ is the greatest subgraph of the underlying graph of $\eqgraph{G}$ on which $f$ and $g$ are equal, together with the inclusion $e: \eqgraph{E \rightarrow G}$. The equivalence relation is preserved ($\sim_E$ is $\sim_G$ restricted to the set of vertices of $E$).

@@

    %%% IDEA: far vedere che, poiché i colimiti in una categoria di prefasci sono calcolati punto punto, allora ho un pushout su ogni componente (E, V, Q). Dalla prop nella parte sull'adesività si può concludere che tutti e tre i pushout preservano mono e quindi mono è preservata su tutte e tre le componenti, quindi reg  mono.
\end{proof}

\begin{cor}\label{cor:quot_pres_reg_mono_pushouts}
    Quotient functor preserves pushouts along regular monomorphisms.
\end{cor}

}\color{black}
\todo[color=green!40]{
COSE DA FARE
>Quoziente crea i limiti (questo dovrebbe derivare direttamente dalla sezione precedente ) 
>Quoziente crea certi pushout (questo va fatto ex novo)
}
% Regular monos in $\mathbf{EqGrph}$ are then natural transformations $\eta: \eqgraph{G \dot\rightarrow H}$ such that $\eta_Q$ is mono (\Cref{prop:reg_mono_in_EG_are_mono_in_graph}). Pullbacks preserve regular monomorphisms (\Cref{prop:monos_pres_by_pullback}). Regular monos are preserved also by pushouts (\Cref{lemma:reg_mono_pres_by_pushouts}). Again, since (co)limits are computed pointwise, and in $\Set$ monos are preserved by pushouts (since $\Set$ is adhesive), we can conclude what follows.

@@

\begin{proof}
    We will show now that $Q$ reflects pullbacks.
    Consider the following diagram in $\mathbf{Graph}$
    \[
        \begin{tikzcd}
            & Q(\eqgraph{G}) \arrow[d, "Q(f)"] \\
            Q(\eqgraph{H}) \arrow[r, "Q(g)"swap] & Q(\eqgraph{K})
        \end{tikzcd}
    \]
    and let $(\mathcal P, p_{Q(f)}, p_{Q(g)})$ be the pullback. We want to show that, if  $(\eqgraph{P}, p_f, p_g)$ is the pullback of the arrows $f$ and $g$ in $\mathbf{EqGrph}$, then $Q(\eqgraph{P}) = \mathcal{P}$, $Q(p_f) = p_{Q(f)}$ and $Q(p_g) = p_{Q(g)}$. 
    \color{green}{
        from \Cref{prop:limits_in_EqGrph}, we note that a limit in $\mathbf{EqGrph}$ is nothing more than the limit graph embedded with an equivalence relation.
        % IDEA: come si calcola un pullback in EgGrph? --> componente per componente, e poi su Q la least equiv relation such that ..., quindi applicando il funtore quoziente ottengo esattamente il pullback dei quozienti.
    }\color{black}
    Hence, from \Cref{cor:quot_preserves_co_lim}, we can conclude that $Q$ creates pullbacks.
\end{proof}

@@

% \begin{proof}
%     Consider the following situation
% \end{proof}
